---
title: Datasets
author: Ramnath Vaidyanathan
---

```{r setup}
# Load packages ----
library(tidyverse)
library(rvest)
library(tidyr)
library(readr)
library(usethis)
```


## Baby Names

The data (name, year of birth, sex, and number) are from a 100 percent sample of
Social Security card applications for 1880 onward.

[Reference](https://catalog.data.gov/dataset/baby-names-from-social-security-card-applications-national-level-data)

```{r}
url_babynames <- "https://www.ssa.gov/oact/babynames/names.zip"
tf <- tempfile(fileext = '.zip')
download.file(url_babynames, destfile = tf)
td <- tempdir()
unzip(tf, exdir = td)
babynames <- fs::dir_ls(td, glob = '*.txt') %>%
  purrr::map(~ {
    d <- readr::read_csv(
      .x, 
      col_names = c('name', 'sex', 'n'),
      col_types = list(col_character(), col_character(), col_number())
    )
    d$sex <- as.character(d$sex)
    d$year <- basename(.x) %>% 
      tools::file_path_sans_ext() %>% 
      gsub('yob', '', .)%>% 
      as.numeric()
    d
  }) %>% 
  do.call(bind_rows, .)
babynames
# write.csv(babynames, file = gzfile('babynames.csv.gz'), row.names = FALSE)
```



## Social Security Applications

This table shows the number of applicants for a Social Security card by year of
birth and sex. The number of such applicants is restricted to U. S. births where
the year of birth, sex, State of birth (50 States and District of Columbia) are
known, and where the given name is at least 2 characters long.

[Reference](https://www.ssa.gov/oact/babynames/numberUSbirths.html)

```{r}
# SOURCE: https://github.com/hadley/babynames/blob/master/data-raw/applicants.R

page <- read_html("https://www.ssa.gov/oact/babynames/numberUSbirths.html")

ssa <- page %>% html_nodes("table") %>% .[[1]] %>% html_table() %>% tbl_df()
names(ssa) <- c("year", "M", "F", "total")
ssa$total <- NULL

ssa$M <- parse_number(ssa$M)
ssa$F <- parse_number(ssa$F)

applicants <- ssa %>%
  gather(sex, n_all, M:F) %>%
  arrange(year, sex) %>%
  mutate(n_all = as.integer(n_all)) %>%
  arrange(year, sex)

applicants
# write_csv(applicants, "applicants.csv")
```
## Life Expectancy Tables

